Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change

### 2025-01-01 ms.qmd
Dealt with R1 and R2 comments in introduction and fig 1.

Reversible phenotypic plasticity, such as physiological acclimation, is relatively rapid and can be fine-tuned to environmental conditions making it the first ‘line-of-defense’ against environmental change [@Dewitt1998]. For example, physiological rates are known to speed up as temperature increases because of the thermodynamic effects on chemical reaction rates -- so called 'acute' temperature responses (@fig-concept). However, longer-lasting (days-weeks) temperature increases that move environmental conditions away from thermal maxima of physiological rate functions (i.e., thermal performance curves) can be mitigated by acclimation, which adjust reaction rates [@Seebacher2015-wi; @Havird2020-mj]. Physiological acclimation is driven by endocrine and epigenetic processes that change the underlying physiology to allow organisms to maintain physiological performance around a fitness optimum despite changes in the environment [@seebacher2019histone; @little2013thyroid; @taff2016endocrine]. Acclimation therefore alters acute thermal sensitivity to offset the potentially negative effects of acute temperature changes (e.g., higher energetic demands). Acclimation, however, does not necessarily result in complete compensation in response to environmental change [ *sensu* @huey1999testing]. Rather, increased physiological rates are often only partially compensated such that ectotherms acclimated to, and measured at, warmer temperatures (R2.2 in @fig-concept) have higher physiological rates than those acclimated to, and measured at, cooler temperatures (R1.1 in @fig-concept) [@huey1999testing; @Havird2020-mj].

>>>>

Reversible phenotypic plasticity, such as physiological acclimation, is relatively rapid and can be fine-tuned to environmental conditions making it the first ‘line-of-defense’ against environmental change [@Dewitt1998]. For example, physiological rates are known to speed up as temperature increases because of the thermodynamic effects on chemical reaction rates -- so called 'acute' temperature responses. However, longer-lasting (days-weeks) temperature increases that move environmental conditions away from thermal optima can be mitigated by acclimation, which adjust reaction rates or the thermal optima itself [@Seebacher2015-wi; @Havird2020-mj]. Physiological acclimation is driven by endocrine and epigenetic processes that change the underlying physiology to allow organisms to maintain physiological performance around a fitness optimum despite changes in the environment [@seebacher2019histone; @little2013thyroid; @taff2016endocrine]. Acclimation therefore alters acute thermal sensitivity to offset the potentially negative effects of acute temperature changes (e.g., higher energetic demands). Acclimation, however, does not necessarily result in complete compensation in response to environmental change [ *sensu* @huey1999testing]. Rather, increased physiological rates are often only partially compensated such that ectotherms acclimated to, and measured at, warmer temperatures have higher physiological rates than those acclimated to, and measured at, cooler temperatures [@huey1999testing; @Havird2020-mj].
